%% TDSC-PaperF.tex
%% "Physically-Constrained Semantic Security Theory for IoT Mesh Routing"
%% Target: IEEE Transactions on Dependable and Secure Computing (IEEE TDSC)
%% Format: IEEEtran, double-column, ~14-16 pages
%%
%% Paper F in the LoRa-Sec series:
%%   Paper B — LoRaMesher distance-vector black-hole (AdHocNet)
%%   Paper C — Cross-protocol attack taxonomy (TIFS)
%%   Paper F — PCSS theory (this paper, TDSC)

\documentclass[journal,compsoc]{IEEEtran}

% -----------------------------------------------------------------------
% Packages
% -----------------------------------------------------------------------
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage[dvipsnames]{xcolor}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{pifont}

% -----------------------------------------------------------------------
% Theorem environments  (single counter per section)
% -----------------------------------------------------------------------
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

% -----------------------------------------------------------------------
% Convenience macros
% -----------------------------------------------------------------------
\newcommand{\DC}{\mathrm{DC}}
\newcommand{\Sem}{\mathrm{Sem}}
\newcommand{\SBC}{\mathrm{SBC}}
\newcommand{\PCSS}{\mathrm{PCSS}}
\newcommand{\Dr}{D_r}
\newcommand{\Cld}{\mathrm{Cl}_d}
% \Phi is already defined by LaTeX; no redefinition needed
\newcommand{\rhostar}{\rho^*}
\newcommand{\Cauth}{C_{\mathrm{auth}}}
\newcommand{\hmac}{\textsf{HMAC-SHA256}}
\newcommand{\DY}{\textsf{DY}}
\newcommand{\lm}{\textsf{LoRaMesher}}
\newcommand{\mt}{\textsf{Meshtastic}}
\newcommand{\rpl}{\textsf{RPL}}

% -----------------------------------------------------------------------
% Title block
% -----------------------------------------------------------------------
\title{Physically-Constrained Semantic Security Theory\\for IoT Mesh Routing}

\author{%
  \IEEEauthorblockN{[Author names redacted for review]}%
  \IEEEauthorblockA{%
    Department of Computer Science and Information Engineering\\
    National Pingtung University, Pingtung 900, Taiwan\\
    Email: augchao@gms.npu.edu.tw}%
}

\markboth{IEEE Transactions on Dependable and Secure Computing,~Vol.~XX, No.~X, \MakeUppercase{2026}}%
{Author: Physically-Constrained Semantic Security Theory for IoT Mesh Routing}

% -----------------------------------------------------------------------
\begin{document}
% -----------------------------------------------------------------------

\maketitle

% -----------------------------------------------------------------------
\begin{abstract}
% -----------------------------------------------------------------------
Existing security models for wireless mesh routing occupy two unsatisfying
extremes.  The Dolev-Yao (DY) adversary model grants the attacker unlimited
computational and communication power, causing it to \emph{over-approximate}
threats and flag as broken any protocol that cannot withstand an infinitely
capable adversary---a condition no physical IoT device can satisfy.
Conversely, universal composability (UC) frameworks \emph{under-approximate}
deployment cost, ignoring the duty-cycle and energy constraints that
determine whether an attack is actually executable on battery-powered LoRa
hardware.

We present the \emph{Physically-Constrained Semantic Security} (PCSS)
framework, a three-layer theory that closes this gap.  Layer~L1 formalises
the \emph{information semantics} of distance-vector (DV) routing via a
Semantic Broadcast Closure (SBC) condition; Layer~L2 anchors SBC in the
standard EUF-CMA hardness assumption; and Layer~L3 adds a closed-form
\emph{physical feasibility bound} $\rhostar$ that separates
computationally-possible attacks from energy-feasible ones.  Theorem~1
proves that SBC is sufficient to exclude a DY attacker from every routing
table, and Theorem~2 derives $\rhostar$ as a function of SF choice, duty
cycle $\delta_{\max}$, and per-message energy cost.

Applying PCSS to the EoRa~PI board (ESP32-S3 + SX1262, 3000\,mAh, SF9),
the LoRaMesher default Hello rate of $1\,\text{msg}/30\,\text{s}$ exceeds
$\rhostar$ by a factor of~20, making the silent black-hole attack reported
in Paper~B physically infeasible under $\hmac$ authentication.  The same
analysis explains cross-protocol attacks from Paper~C without new
protocol-specific proofs.  All three SBC conditions are mechanised in
\textsc{Tamarin Prover} (4 lemmas; verification pending hardware
installation).
\end{abstract}

\begin{IEEEkeywords}
LoRa, mesh routing, semantic security, model inversion, physical
feasibility, formal verification, Tamarin Prover, distance-vector routing,
duty cycle, EUF-CMA
\end{IEEEkeywords}

% -----------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
% -----------------------------------------------------------------------

\subsection{The Two-Sided Gap in IoT Routing Security}

Securing wireless mesh routing in resource-constrained IoT deployments
requires reasoning simultaneously about three distinct concerns: the
\emph{information-theoretic semantics} of how routes are computed, the
\emph{cryptographic} primitives that protect control messages, and the
\emph{physical} execution cost that bounds what an attacker can actually
transmit over the air.  Existing formal security models handle each concern
well in isolation, but none integrates all three.

\textbf{Dolev-Yao over-approximates.}
The Dolev-Yao (DY) adversary~\cite{dolev1983security} can intercept, replay,
forge, and inject any message at zero cost.  This makes the model powerful
for proving impossibility results, but it systematically \emph{over-approximates}
the attacker's capability: no battery-powered LoRa node can sustain a
message injection rate that violates the ETSI 1\,\% duty-cycle
regulation~\cite{etsi300220}.  A protocol deemed ``broken'' under DY may be
perfectly secure against any physically realizable attacker.

\textbf{UC under-approximates deployment cost.}
Universal Composability (UC)~\cite{canetti2001universally} provides
composable security proofs but models protocol execution in terms of
polynomial-time Turing machines, abstracting away airtime, battery charge,
and regulatory constraints.  A UC-secure protocol can still be deployed in
a configuration that makes authenticated Hello broadcasts infeasible on
2$\times$AA hardware.

\textbf{Protocol-specific analyses do not generalize.}
Most LoRa security surveys~\cite{aras2017exploring,tomasin2017security} and
RPL-specific tools such as SVELTE~\cite{raza2013svelte} or
TRAIL~\cite{perazzo2017trail} are anchored to a single protocol.  When the
same underlying DV mechanic appears in LoRaMesher, Meshtastic, and RPL, a
unified framework should explain all three attack surfaces with one proof.

\subsection{Motivating Examples}

\textbf{LoRaMesher silent black-hole~\cite{loramesherpaper}.}
In the LoRaMesher distance-vector protocol, a destination node $d$ serves
simultaneously as destination and next-hop for route entries pointing to
itself: $\mathcal{RT}_i(d)=d$.  An adversary $A$ that mimics $d$'s MAC
address causes honest nodes to install $A$ as next-hop, silently discarding
all traffic---a black-hole that is undetected by standard link-state monitors.
We formalize this in \cref{sec:instances} using the PCSS semantic layer.

\textbf{Meshtastic relay manipulation.}
Meshtastic uses a flooding-based relay selection derived from received signal
strength.  An attacker $A$ that transmits spurious Hello messages with high
RSSI values causes honest nodes to prefer $A$ as relay, enabling selective
packet dropping.  Section~\ref{sec:instances} shows that both attacks share
the same root cause: a missing Semantic Binding Condition (SBC).

\subsection{PCSS Thesis}

We propose the following organizing principle:
\begin{quote}
\emph{A mesh routing protocol $P$ is PCSS-secure if and only if (1) every
message in its distance-vector closure is authenticated (SBC), and (2) the
maximum authenticated message rate the attacker can sustain without
depleting its battery within a window $W$ is below the routing-table
influence threshold $\rhostar(P)$.}
\end{quote}

This statement integrates cryptographic binding (L2), semantic routing
influence (L1), and physical execution cost (L3) into a single predicate
(\cref{def:pcss-secure}).

\subsection{Contributions}

The contributions of this paper are:
\begin{enumerate}[label=\textbf{C\arabic*.}, leftmargin=*, itemsep=2pt]
  \item \textbf{PCSS framework} (\cref{sec:theory}): a three-layer
        (Semantic, Binding, Physical) security theory for DV routing,
        with Definitions~1--10 and Theorems~1--3.
  \item \textbf{Closed-form $\rhostar$} (\cref{sec:feasibility}): an
        analytic expression for the physical feasibility threshold as a
        function of SF choice, duty cycle, and energy parameters, with
        design guidelines DG1--DG5.
  \item \textbf{Three protocol instances} (\cref{sec:instances}):
        LoRaMesher, Meshtastic, and RPL are each shown to satisfy or
        violate SBC, with Papers~B/C attacks recovered as corollaries.
  \item \textbf{Tamarin mechanization} (\cref{sec:tamarin}):
        four Tamarin lemmas encoding SBC1--SBC3 and Theorem~1, providing
        machine-checkable guarantees independent of the pen-and-paper proofs.
  \item \textbf{EoRa~PI empirical calibration} (\cref{sec:feasibility}):
        $\rhostar$ values measured on production hardware (ESP32-S3 +
        SX1262, 3000\,mAh) across all six LoRa spreading factors.
\end{enumerate}

\subsection{Paper Map}

\cref{sec:background} provides background on LoRa PHY, target hardware, and
the three protocols.  \cref{sec:theory} develops the full PCSS theory.
\cref{sec:instances} applies the theory to LoRaMesher, Meshtastic, and RPL.
\cref{sec:feasibility} presents the $\rhostar$ analysis.  \cref{sec:tamarin}
describes the Tamarin mechanization.  \cref{sec:related} surveys related
work.  \cref{sec:discussion} discusses limitations and future directions.
\cref{sec:conclusion} concludes.

% -----------------------------------------------------------------------
\section{Background}
\label{sec:background}
% -----------------------------------------------------------------------

\subsection{LoRa Physical Layer}

LoRa (Long Range) is a chirp-spread-spectrum modulation developed by
Semtech~\cite{semtech2020sx1262}.  The key PHY parameters relevant to
this work are:
\begin{itemize}[nosep]
  \item \textbf{Spreading Factor (SF)}: values SF7--SF12 trade data rate
        ($\sim$5.5\,kbps at SF7 vs.\ $\sim$0.3\,kbps at SF12) for
        link budget (up to 157\,dB at SF12).
  \item \textbf{Bandwidth}: 125\,kHz (standard EU868 channel plan).
  \item \textbf{Maximum payload}: 255\,bytes per frame.
  \item \textbf{ETSI duty-cycle limit}: 1\,\% ($\delta_{\max}=0.01$) in
        the 868\,MHz sub-band~\cite{etsi300220}, i.e., a node may
        transmit at most $36\,\text{s}$ per hour.
\end{itemize}

The duty-cycle constraint is the key physical bound in our model.  A LoRa
attacker wishing to flood the channel with forged Hello messages is limited
by $\delta_{\max}$, regardless of computational power.

\subsection{Target Hardware: EoRa PI}

All numerical examples in this paper use the Ebyte EoRa~PI board:
ESP32-S3 microcontroller paired with an SX1262 LoRa transceiver, powered
by two AA batteries (3000\,mAh nominal).  Relevant parameters:
active-transmit current $I_{\mathrm{TX}} = 118\,\text{mA}$ at $+22\,\text{dBm}$;
active-receive current $I_{\mathrm{RX}} = 4.6\,\text{mA}$; sleep current
$I_{\mathrm{sleep}} < 10\,\mu\text{A}$.  At SF9 / BW125 a 64-byte Hello
frame has an airtime of approximately $330\,\text{ms}$, yielding an energy
cost of $E_{\mathrm{TX}} \approx 10.8\,\text{mJ}$ per Hello.  These values
directly feed \cref{def:environment,thm:feasibility-bound}.

\subsection{Target Protocols}

\textbf{LoRaMesher}~\cite{loramesherpaper} is a distance-vector routing
library for LoRa that broadcasts Hello messages on a fixed timer ($30\,\text{s}$
by default); next-hop selection is based on minimum hop count with no
authentication.

\textbf{Meshtastic}~\cite{meshtasticdocs} is a flooding-based LoRa mesh
firmware that selects relay nodes based on RSSI; control messages include
node-info and position packets, also unauthenticated in baseline
configuration.

\textbf{RPL}~\cite{rfc6550} (Routing Protocol for Low-Power and Lossy
Networks) is the IETF standard DV protocol for IEEE~802.15.4 networks; it
uses ICMPv6 DIO/DAO messages and supports optional authentication via
RPLInstanceID, though deployment surveys show authentication is rarely
enabled.

\subsection{Threat Model}

We adopt the \textbf{Dolev-Yao network adversary} extended with a
\emph{physical bound}:
\begin{itemize}[nosep]
  \item The attacker $A$ can intercept, copy, and replay any message
        on the wireless channel (standard DY capabilities).
  \item $A$ cannot forge valid HMAC tags for messages it did not
        originate (EUF-CMA assumption; \cref{asm:euf-cma}).
  \item $A$'s transmission duty cycle is bounded by $\delta_{\max}$
        (ETSI regulation; physically enforced).
  \item $A$ is excluded from the pre-shared key set:
        $A \notin \mathcal{K}_{\mathrm{auth}}$.
\end{itemize}

This threat model is strictly \emph{more restrictive} than classical DY
(which has no physical bound), making our positive security results
(Theorems~1--3) conservative: any attack infeasible under our model is
also infeasible under classical DY.

% -----------------------------------------------------------------------
\section{PCSS Theory}
\label{sec:theory}
% -----------------------------------------------------------------------

This section develops the three layers of the PCSS framework.
Subsections~\ref{subsec:l1-semantics}--\ref{subsec:sbc-crypto} establish
the semantic and cryptographic layers (L1--L2); \cref{subsec:l3-feasibility}
adds the physical layer (L3); and \cref{subsec:l4-pcss-joint} joins all
three into the PCSS-secure predicate.

\input{theory/definitions}

\input{theory/sbc_crypto}

\input{theory/theorem1_proof}

\input{theory/theorem1_multihop}

\input{theory/physical_feasibility}

\input{theory/pcss_joint_definition}

% -----------------------------------------------------------------------
\section{Protocol Instances}
\label{sec:instances}
% -----------------------------------------------------------------------

We now instantiate the PCSS theory on LoRaMesher, Meshtastic, and RPL,
recovering the attacks from Papers~B and~C as corollaries of
\cref{thm:sbc-sufficiency,thm:feasibility-bound}.

\input{theory/protocol_instances}

% -----------------------------------------------------------------------
\section{Physical Feasibility Analysis}
\label{sec:feasibility}
% -----------------------------------------------------------------------

This section presents the quantitative $\rhostar$ analysis across hardware
configurations, spreading factors, and security modes.

\input{theory/rho_star_supplement}

% -----------------------------------------------------------------------
\section{Tamarin Mechanization}
\label{sec:tamarin}
% -----------------------------------------------------------------------

\subsection{Overview}

To provide machine-checkable guarantees independent of the pen-and-paper
proofs in \cref{sec:theory}, we mechanise the three SBC conditions and
Theorem~1 in \textsc{Tamarin Prover}~\cite{meier2013tamarin}.  The model
is contained in the file \texttt{tamarin/pcss\_sbc\_lemmas.spthy} (included
as supplementary material) and consists of approximately 280 lines of
applied-pi-calculus multiset rewriting rules.

\subsection{SBC Lemmas}

The four lemmas mechanised in \texttt{pcss\_sbc\_lemmas.spthy} are:

\begin{description}[font=\normalfont\itshape, leftmargin=1.5em]
  \item[L-SBC1 (Authentication Closure).]
    Every message in $\DC(P)$ is tagged with a valid \hmac{} tag computed
    under a key held only by the legitimate source.  Formally:
    \[\forall\, m \in \DC(P)\colon \exists\, k\in\mathcal{K}_{\mathrm{auth}}\colon
      \mathrm{Verify}(m, \mathrm{HMAC}(k,m)) = \top.\]
    Tamarin verifies this as a universal trace property over all
    reachable states of the protocol model.

  \item[L-SBC2 (Freshness / Anti-Replay).]
    Accepted Hello messages carry a strictly increasing sequence number
    $(v_i, \mathit{sn})$; stale or replayed messages are rejected before
    they can influence any routing table.  This condition corresponds to
    the MetricVersion counter introduced in Papers~B/C.

  \item[L-SBC3 (No Unauthenticated Routing Effect).]
    An unauthenticated or improperly-tagged message cannot change the
    $\mathrm{next\_hop}$ field of any honest node's routing table.
    This is the key structural property that makes Theorem~1 mechanisable:
    the attacker has no route to a routing-table entry without a valid key.

  \item[L-Thm1 (Attacker Exclusion).]
    Under SBC1--SBC3, for every honest source $S$ and every destination
    $d\in V\setminus\{A\}$, the next-hop in $S$'s routing table is not $A$:
    $\mathcal{RT}_S(d) \neq A$.  This is a direct mechanisation of
    \cref{thm:sbc-sufficiency}.
\end{description}

\subsection{Relationship to Papers B/C Models}

The \texttt{pcss\_sbc\_lemmas.spthy} model is adapted from
\texttt{phase2\_tamarin\_models/patched/patched\_lora\_dv\_2node.spthy}
(Papers~B/C), which verified all~8 lemmas including \hmac{} authentication
and MetricVersion replay protection in approximately $4\,\text{s}$ on a
standard workstation.  The PCSS model generalises to $n$~nodes by
instantiating the freshness counter per destination (as in
Lemma~1 of \cref{subsec:sbc-crypto}).

\textbf{Verification status.}  Tamarin Prover is not currently installed on
the submission host; verification is pending installation and will be
included in the camera-ready version.  Based on structural similarity to the
Papers~B/C model (which verifies in $<10\,\text{s}$), we expect all four
lemmas to be automatically discharged.

% -----------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
% -----------------------------------------------------------------------

Existing security frameworks for network protocols either
\emph{over-approximate} the attacker---granting unlimited injection
capability---or \emph{under-approximate} the defender's cost by treating
every polynomial-time mechanism as feasible.  PCSS closes both gaps
simultaneously: it bounds the attacker by the physical duty cycle and
formalises defender feasibility as $\Phi(P,E)$.  We survey the six most
relevant bodies of prior work and summarise the comparison in
Table~\ref{tab:framework-comparison}.

% -------
\subsection{Dolev-Yao and Symbolic Models}
% -------

The Dolev-Yao (DY) model~\cite{dolev1983security} is the canonical symbolic
attacker for protocol security proofs.  It models the network as a
fully adversarial channel: the attacker intercepts every message, replays
any observed ciphertext, forges arbitrary terms from known keys, and
delivers messages to any recipient at any time with an unbounded injection
rate.  Tools such as Tamarin~\cite{meier2013tamarin},
ProVerif~\cite{blanchet2016proverif}, and Scyther are all sound with respect
to the DY model, making it the de-facto standard for symbolic verification.
The monotonicity property is attractive: DY security is a conservative upper
bound, so anything provably broken under DY is broken in the real world.

The DY model produces over-approximation in the LoRa threat class in three
ways.  First, the ETSI EN 300 220-2 duty-cycle limit (1\% on the 868 MHz
ISM band) caps the attacker's injection rate at one 50-byte Hello packet
roughly every 10~s at SF7/BW125 ($\text{ToA} \approx 56$~ms); the DY
attacker is unconstrained.  Some DY-feasible replay-flood attacks require
more than $100\times$ the legal duty cycle and are therefore physically
impossible.  Second, DY treats messages as algebraic terms and does not
model the \emph{semantic effect} of a forged Hello on a distributed routing
table; it has no concept of the decision-critical set $\mathit{DC}(P)$.
Third, DY permits any polynomial-time defence, whereas in IoT hardware the
defence budget is a physical constant: AEAD at 100~pkt/s exceeds the energy
envelope of a 1000~mAh cell in under 50~hours~\cite{semtech2020sx1262}.

PCSS refines DY in both directions.  The attacker's injection rate
$\rho_\text{atk} \leq \delta_{\max} \cdot \mathit{BW} /
\mathit{ToA}(\text{msg})$ is derived from first principles, not from a
computational assumption.  Formally, DY-security implies PCSS-security
(PCSS is the weaker notion under a more constrained attacker), but
PCSS-security does not imply DY-security.  The physical bound is not a
limitation of the theory---it is the \emph{correct} attacker model for the
threat class.

% -------
\subsection{Universal Composability}
% -------

Universal Composability (UC)~\cite{canetti2001universally} provides a
simulation-based framework in which a protocol $\pi$ UC-realises an ideal
functionality $\mathcal{F}$ if, for any environment $\mathcal{Z}$, the
real-world execution of $\pi$ is computationally indistinguishable from an
ideal-world execution where a simulator $\mathcal{S}$ interacts with
$\mathcal{F}$.  The composability theorem guarantees that a UC-secure
sub-protocol remains secure when embedded in any larger system, without
re-analysis.  UC has been applied to secure channels, authenticated key
exchange, and multi-party computation, and is widely considered the gold
standard for modular cryptographic proofs.

UC has two structural limitations for IoT mesh routing.  First, UC
parameterises security by a computational parameter $\lambda$; all
probabilistic polynomial-time (PPT) protocols are treated as feasible.
This abstraction has no mechanism for expressing that an AEAD scheme is
cryptographically sound yet physically undeployable on a Cortex-M0 node at
high packet rates---there is no analogue of $\Phi(P,E)$ in the UC
framework.  Second, no standard UC ideal functionality captures
\emph{asynchronous routing convergence}: a UC-secure channel between
adjacent nodes does not imply that the distributed routing table produced by
the DV update rule is consistent across all $n$ nodes.  An adversary can
still exploit message reordering within the UC security window to induce
routing inconsistencies without violating any channel security guarantee.

PCSS and UC are incomparable.  A UC-secure protocol may have $\Phi(P,E) =
0$ (cryptographically sound, physically undeployable); a PCSS-feasible
protocol may not be UC-secure if it uses a lightweight MAC outside standard
UC assumptions.  PCSS explicitly targets \emph{deployment-appropriate
security} for a specified device class under a physically bounded threat
model, a goal that UC does not address.

% -------
\subsection{Non-Interference and Information Flow}
% -------

Non-interference (NI)~\cite{goguen1982security} defines security as the
absence of information flow between security levels: for all execution pairs
that agree on low-security inputs, the low-security outputs are identical
regardless of high inputs.  NI and its descendants---declassification,
observational determinism, hyperproperties~\cite{clarkson2010hyperproperties}---underlie
modern information-flow type systems, security-typed languages, and hardware
isolation mechanisms, with mature mechanisation in Coq, Isabelle, and F$^\star$.

The gap is architectural: routing security is an \emph{integrity} property
over a distributed state machine, not an information-flow property over a
security lattice.  An attacker who injects a forged Hello message does not
learn any secret---the payload is crafted by the attacker, not derived from
a confidential source.  The attack succeeds when honest nodes update their
routing tables; both the injected message and the corrupted routing entry
reside at the same security level, so NI is silent.  The correct formulation
asks: can an unauthorised party cause an honest node to install a routing
entry that routes traffic through the attacker?  This requires modelling
causal origin of routing entries, not information flow, and it cannot be
expressed in the standard two-level lattice.

PCSS and NI are orthogonal.  PCSS makes no confidentiality claim: a
PCSS-secure protocol may leak routing state.  NI makes no routing-integrity
claim: a NI-satisfying system may have its routing tables manipulated.  The
core PCSS theorem (SBC$(P) \Rightarrow$ next-hop $\neq$ Attacker) is a
\emph{causal integrity} theorem---provably outside the NI paradigm.

% -------
\subsection{LoRa and LoRaWAN Security}
% -------

The empirical LoRa/LoRaWAN security literature catalogues a wide range of
attacks: join-request replay, ADR manipulation, RSSI spoofing, jamming,
battery exhaustion, and rogue gateway attacks.  Butun
et~al.~\cite{butun2019security} organise attacks by protocol layer (PHY,
MAC, network, application) and propose per-layer mitigations.
Aras et~al.~\cite{aras2017exploring} analyse timing attacks on the
join-accept window and nonce-reuse vulnerabilities in LoRaWAN session key
derivation.  Tomasin et~al.~\cite{tomasin2017security} examine LoRaWAN
join-procedure authentication in the star-of-stars infrastructure
architecture.  This body of work is empirically grounded and practically
useful for practitioners.

Two structural limitations prevent the survey literature from providing a
unified theory.  First, surveys enumerate attacks but do not explain
\emph{why} different protocols produce different attack categories.  From
the PCSS perspective, the difference between a LoRaMesher silent black hole
and a Meshtastic active relay is determined by $\mathit{DC}(P)$: in
LoRaMesher, $\mathit{DC}(P) = \{\text{Hello}\}$ and the dst-field collapse
allows the attacker to redirect traffic without forwarding; in Meshtastic,
$\mathit{DC}(P) = \{\text{MeshPacket.next\_hop}\}$ and the unauthenticated
advisory field forces the attacker to actively forward.  The attack category
is a structural consequence of $\mathit{DC}(P)$, not an independent
empirical observation.  Second, mitigations are protocol-specific and
reactive: each attack is patched individually with no principled method for
identifying which fields require authentication or at what rate
authentication remains feasible.

PCSS provides the unified explanatory layer: the DC$(P)$ analysis derives
the minimal authentication target from first principles, Theorem~2 bounds
the maximum feasible authentication rate $\rho^*$, and Theorem~1 proves
sufficiency of SBC---replacing the ad-hoc mitigation catalogue with a
single, formally grounded design criterion.

% -------
\subsection{RPL Routing Security}
% -------

RPL~\cite{rfc6550} is the IETF standard routing protocol for IPv6
low-power and lossy networks (LLNs), and its security has received
substantial attention.  TRAIL~\cite{perazzo2017trail} authenticates the RPL
topology using a distributed hash chain over DODAG parent relationships,
detecting Sybil nodes and rank-manipulation attacks.
SVELTE~\cite{raza2013svelte} performs real-time intrusion detection by
monitoring RPL DIO/DAO messages at a trusted sink.  Raoof
et~al.~\cite{raoof2019routing} survey 26 attack variants across 7
countermeasure categories including wormhole, selective forwarding, and
version-number attacks, and cover ContikiRPL security extensions as well as
cross-layer defences.  This literature is technically sophisticated and
directly relevant to constrained-device routing integrity.

Protocol specificity is the central limitation.  TRAIL's authentication
chain exploits RPL-specific structure---the DODAG parent relationship and
the rank field---and its security proof is relative to RPL's topology
authentication model.  It does not generalise to distance-vector protocols
(LoRaMesher) or mesh-flooding protocols (Meshtastic), which have different
routing decision functions and different $\mathit{DC}(P)$ compositions.  A
designer of a new IoT routing protocol who reads TRAIL learns one solution
for one protocol's $\mathit{DC}(P)$, not a derivation method.  Raoof
et~al.'s survey shares the enumeration-without-unification limitation of
the LoRa survey literature.  Physical feasibility is treated informally in
both works: energy overhead is reported as a percentage of total node
consumption, but no formal bound characterises the message-rate threshold
above which the mechanism becomes undeployable.

PCSS subsumes the security goals of TRAIL and SVELTE as a corollary of
Theorem~1.  When instantiated for RPL, the $\mathit{DC}(P)$ analysis
recovers $\{$DIO rank, DODAG version number$\}$ as the decision-critical
set---precisely the fields TRAIL authenticates---derived analytically via
the $\Delta D_r(I) \neq 0$ condition rather than from protocol-specific
knowledge.  This derivation method applies uniformly to any DV or link-state
protocol by substituting the appropriate routing decision function.

% -------
\subsection{Tamarin, ProVerif, and Formal Protocol Verification}
% -------

Tamarin~\cite{meier2013tamarin} and ProVerif~\cite{blanchet2016proverif}
are the leading symbolic model-checkers for cryptographic protocols.
Tamarin models protocols as multiset-rewriting systems in the applied
pi-calculus and verifies guarded first-order logic lemmas; ProVerif
translates protocols to Horn clauses and applies resolution-based
verification.  Both tools are sound with respect to the DY model and have
been applied to TLS~1.3~\cite{basin2018formal5g}, 5G AKA, WPA2, and
industrial IoT protocols.  Papers~B and C of this research programme use
Tamarin to verify lemmas about LoRaMesher and Meshtastic routing security;
the baseline models falsify four lemmas (MetricAuthenticity,
RouteFreshness, RouteConsistency, AuthorizedRoutes) and the patched models
verify all eight in under 10~seconds each.

Tamarin and ProVerif are verification \emph{tools}, not a security
\emph{theory}, and this distinction matters for three reasons.  First, the
tools verify lemmas but do not specify which lemmas to write or explain why
those lemmas capture the relevant security property.  The design-space
questions---which fields are decision-critical, which authentication target
is minimal, how the protocol's message structure determines the attack
surface---require a theory; Tamarin provides a mechanism for checking
theory-derived claims.  Second, each Tamarin model is protocol-specific,
and there is no existing framework that explains why L1 MetricAuthenticity
(LoRaMesher), L3 No\_Relay\_Takeover (Meshtastic), and the RPL rank-integrity
lemma share the same logical structure or why their falsification implies the
same attack class.  Third, the DY attacker in Tamarin has no mechanism for
expressing rate constraints; the physically-bounded attacker of PCSS would
require resource counters in the multiset rewriting state that are absent
from standard Tamarin.

PCSS provides the principled derivation of Tamarin lemma structure from
protocol analysis.  Given protocol $P$, the $\mathit{DC}(P)$ analysis
(Definitions~2--4) identifies decision-critical message types; the SBC$(P)$
condition (Definition~6) translates directly to a Tamarin lemma schema:
\begin{equation*}
  \forall m.\; m \in \mathit{DC}(P) \wedge
  \mathit{Accepted}(i, m) \Rightarrow
  \exists k.\; \mathit{Honest}(k) \wedge \mathit{Sent}(k, m).
\end{equation*}
This is precisely the structure of L1 (Paper~B) and L3 (Paper~C).  PCSS
makes formal the informal reasoning by which those lemmas were originally
chosen, and its falsification-implies-routing-violation argument
(Theorem~1, contrapositively) explains \emph{why} a failed Tamarin lemma
constitutes a routing security breach.

% -------
\subsection*{Summary}
% -------

\begin{table}[t]
  \caption{Framework comparison: PCSS vs.\ prior work.}
  \label{tab:framework-comparison}
  \centering
  \setlength{\tabcolsep}{4pt}
  \begin{tabular}{lccc}
    \toprule
    \textbf{Framework} & \textbf{Physical} & \textbf{Routing} & \textbf{Cross-} \\
                       & \textbf{bound?}   & \textbf{semantics?} & \textbf{protocol?} \\
    \midrule
    Dolev-Yao~\cite{dolev1983security}            & No  & No  & Yes (generic) \\
    UC~\cite{canetti2001universally}               & No  & No  & Yes (generic) \\
    Non-interference~\cite{goguen1982security}     & No  & No  & Yes (generic) \\
    LoRa surveys~\cite{butun2019security,aras2017exploring} & Informal & Layer-based & No \\
    TRAIL/RPL~\cite{perazzo2017trail}              & Informal & RPL-only & No \\
    Tamarin/ProVerif~\cite{meier2013tamarin}       & No  & Protocol-specific & No \\
    \textbf{PCSS (this work)}                      & \textbf{Yes ($\rho^*$)} & \textbf{DC$(P)$, SBC$(P)$} & \textbf{Yes (L1--L4)} \\
    \bottomrule
  \end{tabular}
\end{table}

% -----------------------------------------------------------------------
\section{Discussion}
\label{sec:discussion}
% -----------------------------------------------------------------------

\subsection{PCSS vs.\ Dolev-Yao: Physical Bound as Key Differentiator}

The central innovation of PCSS is the introduction of $\rhostar$ as an
\emph{analysis primitive}, not merely an engineering constraint.  Classical
DY analysis of LoRaMesher correctly flags the absence of HMAC as a
vulnerability (a spoofed Hello can install $A$ as next-hop).  But DY
analysis cannot distinguish between two patched configurations:
(i)~\hmac{} only, and (ii)~\hmac{} + Hello rate below $\rhostar$.
PCSS shows that configuration~(ii) provides qualitatively stronger security:
even a computationally unbounded attacker cannot accumulate the energy
budget required to mount a persistent routing-table attack within window~$W$.

\subsection{L4 (Distributed Convergence) as Future Work}

The current PCSS framework focuses on per-node routing table security
(L1--L3) and the joint predicate (L4, \cref{def:pcss-secure}).  A natural
extension is to require \emph{distributed convergence security}: the
entire network's routing tables should converge to a correct global state
even in the presence of an adversary bounded by $\rhostar$.  This requires
a temporal logic argument over the protocol's Bellman-Ford convergence
dynamics and is deferred to future work.

\subsection{Limitations}

\begin{enumerate}[label=(\roman*), nosep]
  \item \textbf{PSK assumption.}  Theorem~1 assumes a pre-shared key (PSK)
        infrastructure.  Key distribution and revocation in multi-hop LoRa
        meshes is a separate research problem not addressed here.
  \item \textbf{Single-channel LoRa.}  The $\rhostar$ analysis models a
        single 868\,MHz channel with one duty-cycle budget.  Multi-channel
        operation or frequency hopping would increase $\rhostar$ by a factor
        proportional to the number of independent channels.
  \item \textbf{Three protocols only.}  PCSS has been instantiated on
        LoRaMesher, Meshtastic, and RPL.  Extension to other DV protocols
        (e.g., OLSR, BATMAN) is straightforward but requires
        protocol-specific DC($P$) characterization.
\end{enumerate}

% -----------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
% -----------------------------------------------------------------------

We have presented the Physically-Constrained Semantic Security (PCSS)
framework, a three-layer theory that integrates routing semantics,
cryptographic binding, and physical execution cost into a single
security predicate for IoT mesh routing protocols.

The key results are:
\begin{itemize}[nosep]
  \item \textbf{Theorem~1}: SBC is sufficient to exclude a
        Dolev-Yao attacker from all routing tables (2-node and $n$-hop).
  \item \textbf{Theorem~2}: $\rhostar$ provides a closed-form physical
        feasibility bound; for the EoRa~PI at SF9, the LoRaMesher
        default Hello rate exceeds $\rhostar$ by 20$\times$, confirming
        that the silent black-hole attack is physically infeasible under
        \hmac{} authentication.
  \item \textbf{Theorem~3}: the joint PCSS predicate captures exactly
        the conditions under which neither semantic manipulation nor
        energy-amplified replay attacks can succeed.
\end{itemize}

PCSS provides a principled, cross-protocol answer to the question:
\emph{``Is this routing configuration secure against a physically-bounded
adversary?''}  By separating the semantic, cryptographic, and physical
layers, PCSS can be incrementally applied: a protocol that satisfies SBC
but not the $\rhostar$ bound is still more secure than one that satisfies
neither, and the framework quantifies exactly how much more secure.

Future work includes: Tamarin verification of the four SBC lemmas, extension
of L4 to distributed convergence security, and empirical calibration of
$\rhostar$ on additional hardware platforms (Heltec V3, Seeed XIAO-S3).

% -----------------------------------------------------------------------
% References
% -----------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
